---
title: "Zika Illumina analysis"
output: 
  html_document: 
    toc: yes
---

```{r, message = FALSE}
library(dplyr)
library(magrittr)
library(ggplot2)
library(tximport)
library(biomaRt)
library(DESeq2)

knitr::opts_chunk$set(cache = TRUE)
options(stringsAsFactors = FALSE)
```

Setting up a clean rmd of the DESeq2 analysis to have a better document to refer back to. 

## Importing Data

Using abundance data from kallisto, run on the paired-end Zika samples, symlinked from another directory

I already ran kallisto on the data, in `kallisto-allrna`, so I can jump right in at the stage of `tximport`

In order to import the data with the correct transcript and gene information, I can download it from biomart


```{r}
mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

transcript2gene <- getBM(attributes = c("ensembl_transcript_id", "ensembl_gene_id", "hgnc_symbol"), mart = mart) %>% 
  # sorting
  arrange(ensembl_gene_id)
```

```{r}
kallistodirs <- Sys.glob("../data/kallisto-allrna/SRR*")
txi <- tximport(files=paste0(kallistodirs, "/abundance.h5"),
                type = "kallisto",
                tx2gene = transcript2gene,
                # countsFromAbundance = "scaledTPM", # not doing this as DESeq requires raw count values
                ignoreTxVersion = TRUE)
```

Let's grab the metadata too:
```{r}
# the Run col matches with what the folder names are
metadata <- read.delim("../data/SraRunTable.txt") %>% filter(Instrument == "Illumina MiSeq")
# making the appropriate factors
metadata$infection_status  %<>% gsub(pattern = " ", replacement = "", x = .) %>%  as.factor()
metadata$Run  %<>% as.factor
```

```{r}
# setting up row and column names for later, also to just keep track of things in the txi object
colnames(txi$counts) <- metadata$Run
rownames(metadata) <- metadata$Run
```


## Making the DESeq object


Importing from the txi object made of the kallisto runs

```{r}
deseq <- DESeqDataSetFromTximport(txi, metadata, ~ infection_status)
```

filtering the object to only have those with counts greater than 1

```{r}
nrow(deseq)
sum(rowSums(counts(deseq)) <=1)

deseq <- deseq[rowSums(counts(deseq)) > 1, ]
nrow(deseq)
```

## Differential Expression analysis

Doing the differential expression analysis with DESeq

```{r}
dds <- DESeq(deseq)
```

Using an accessor function to get the results:

```{r}
dd.results <- results(dds)
dd.results
```

```{r}
mcols(dd.results, use.names = TRUE)
```

Let's get a summary of the results:

```{r}
summary(dd.results)
```

```{r}
# getting the data into a format I can handle:
dd.df <- dd.results %>% as.data.frame
dd.df$ensembl_gene_id <- row.names(dd.df)

dd.tibble <- dd.df %>% left_join(transcript2gene %>% dplyr::select(-ensembl_transcript_id) %>% distinct, .) %>%
  as_tibble() %>%
  filter(!is.na(padj)) %>%
  mutate(neglogp = -log(padj))
```

```{r}
dd.tibble %>% filter(neglogp > 100) %>% arrange(-neglogp)
```

```{r}
dd.tibble %>% filter(abs(log2FoldChange) > 9) %>% arrange(-abs(log2FoldChange))
```






And why not have a volcano plot:



```{r}
# handmade


dd.tibble %>% ggplot(aes(x = log2FoldChange, y = neglogp)) + 
  geom_point(alpha = 1/5) + 
  geom_hline(yintercept = -log(0.1)) +
  labs(y = "-log(padj)")
```


